---
title: "B4BI Pak Choi Seed Data Graphs and Models for paper"
author: "Heather Jenkins"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---
<!-- output: -->
<!--   word_document: -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=3, dpi=600, warning=FALSE, message=FALSE, echo=FALSE)
options(scipen=999)
library(lme4)
library(tidyverse)
library(readxl)
library(janitor)
library(dplyr)
library(car)
library(emmeans)
library(predictmeans)
library(stringr)
library(GGally)
library(gridExtra)
library(pander)
library(kableExtra)
library(DHARMa)
panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'left', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)
options(knitr.kable.NA = '')
```


# Overview

This document contains the final analysis and graphs of the pak choi seed data.

This analysis was performed using  `r version$version.string`. The *tidyverse*
suite of packages was used for data manipulation and visualisation. The
*emmeans* package was used for model predictions. 
The *DHARMa* package was used to help assess model diagnostics and goodness-of-fit.

The predicted 95% confidence intervals and post-hoc pairwise comparison p-values
are adjusted using the FDR correction to account for multiple tests.

# Data used (how it was summarised)

+ The original data is grouped by Year, Farm, Type_of_farm, Simplified_Boundary, Native_new_age, Treatment, Rep.     
+ The mean of the `Seed viable` is taken. This is the mean across all the pods for each rep. Note the number of pods for each Rep was quite variable.  
+ Then the data is grouped by Year, Farm, Type_of_farm, Simplified_Boundary, Native_new_age, Treatment.  
+ The mean seeds viable is calculated, so we get a mean over the Reps.  
+ For each Year, Farm and Simplified_Boundary combination we have a mean seed counts for each treatment where each rep has the same weight in the mean.  
+ If we didn't do the means with the two group_by steps then each pod would have the same weight in the mean.  
+ I've assumed that the number of pods is a feature of the pak choi plants and nothing to do with the Treatments.  

<br>

```{r}
load(file="Cleaned_datasets/pc_seed_data_summarised_cleaned.rda")
  
pc_sum <- pc_sum %>% mutate(Treatment=fct_recode(Treatment, "HP"="Hand pollinated"),
                            Simplified_Boundary=fct_recode(Simplified_Boundary, 
                                        "Old native SNH"="Old native",
                                        "New native SNH"="New native",
                                        "On-farm bare fence"="Bare fence",
                                        "Off-farm bare fence"="Control farm"
                                        )) %>%
  mutate(Mean_seeds_viable_rounded = round(Mean_seeds_viable, 0))

str(pc_sum, vec.len = 3) 
```

We next remove the data with 'one year native' and 'two year native' in 2023-2024 removed. So 'New native' only has

+ One year old in 2021-2022  
+ Two year old in 2022-2023  
+ Three year old in 2023-2024  

Note that is data is all from Arable farms (Garrickfield and Limbrick), and there are only 4 observations removed.

Tables of counts before and after the observations are removed are below:

```{r}
# str(pc_mod)
table(pc_sum$Year, pc_sum$Native_new_age) %>% kable(format = "pandoc")
pc_sum <- pc_sum %>% filter( !( (Native_new_age %in% c("One-year old native", "Two-year old native")) & Year=="2023-2024" ) )
table(pc_sum$Year, pc_sum$Native_new_age) %>% kable(format = "pandoc")

str(pc_sum, vec.len = 3) 
```

# Graph of data

```{r, fig.height=5, fig.width=6}
# table(pc_sum$Simplified_Boundary:pc_sum$Treatment, pc_sum$Year, pc_sum$Type_of_farm)
set.seed(1989)
pc_sum %>% 
  ggplot(aes(x=Treatment, y=Mean_seeds_viable, color=Treatment:Type_of_farm)) + 
  geom_jitter(height=0, width=0.3, size=1, alpha=0.5) + 
  # geom_point(aes(y = mean_viable), size=2, color=1, data = pc_stats) +
  facet_grid(Year~Simplified_Boundary) + theme_bw(base_size=11) +
  labs(x="", y="Mean number of viable seeds", color="") +
  theme(axis.text.x = element_text(angle=20, hjust=1.0, vjust=1.0)) +  
  scale_color_manual(values = c("chocolate",
                                "#21908CFF",
                                "#FDE725FF",
                                '#3B528BFF',
                                "#5DC863FF",
                                "#440154FF")) + 
  theme(
  legend.position = "bottom",
  legend.margin = margin(t = -10),
  legend.spacing.y = unit(0.1, "cm"),
  legend.key.height = unit(0.3, "cm")
) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(colour = "black")
  )
```

# Modelling comments

+ We remove the Closed treatment from the data.
+ There is no data for Arable farms, Off-farm bare fence, 2021-22, Closed.  
+ The Closed is a control and we can see it "worked" in the graph above.  
+ We fit a complicated model initially and then 3 simplified versions.

```{r}
pc_mod <- pc_sum %>% 
  filter(Treatment != "Closed") %>% 
  droplevels()
```

# Model 0: LMM with 4-way interaction (Y x T_of_F x T x S_B))

## Description

We fit linear mixed-effects models to the data. The response is
Mean_seeds_viable. The fixed-effects are Year, Type_of_farm, Treatment and
Simplified_Boundary plus their four-way interaction. There is also a random
effect for Farm. The residuals were inspected visually and the
simulateResiduals() function from the DHARMa package was used. There was no
evidence of non-normality, dispersion or outliers.

## Diagnostics and ANOVA table

```{r, fig.width=9, fig.height=4}
# shapiro.test(pc_mod$Mean_seeds_viable) # okay
# hist(pc_mod$Mean_seeds_viable, breaks=30)
# mod.lm <- lmer(Mean_seeds_viable ~ Type_of_farm*Treatment*Simplified_Boundary + (1|Year/Farm), data=pc_mod)
# The model below is the same
# mod.lm <- lmer(Mean_seeds_viable ~ Type_of_farm*Treatment*Simplified_Boundary + (1|Year) + (1|Year:Farm), data=pc_mod)
# Tried adding Simplified_Buundary and Treatment to the REs eg (1 + Simplified_Boundary|Year:Farm) but got singular boundary.
# mod.lm <- lmer(Mean_seeds_viable ~ Type_of_farm*Treatment*Simplified_Boundary + (1|Year) + (1|Farm), data=pc_mod)
# AIC(mod.lm); BIC(mod.lm)
# mod.lm <- lmer(Mean_seeds_viable ~ Type_of_farm*Treatment*Simplified_Boundary + (1|Year), data=pc_mod)
# AIC(mod.lm); BIC(mod.lm)
# mod.lm <- lmer(Mean_seeds_viable ~ Type_of_farm*Treatment*Simplified_Boundary + (1|Farm), data=pc_mod)
# AIC(mod.lm); BIC(mod.lm)
# mod.lm <- lmer(Mean_seeds_viable ~ Type_of_farm*Treatment*Simplified_Boundary + (1|Farm/Year), data=pc_mod)
# AIC(mod.lm); BIC(mod.lm)
mod.lm <- lmer(Mean_seeds_viable ~ Type_of_farm*Treatment*Simplified_Boundary*Year + (1|Farm), data=pc_mod)
# AIC(mod.lm); BIC(mod.lm)
# residplot(mod.lm, newwd = FALSE)
# summary(mod.lm)
simres <- simulateResiduals(fittedModel = mod.lm)
plot(simres)
# rm(simres)
kable(Anova(mod.lm),digits=6, format = "pandoc")
```

<br>

The four-way and all three-way interactions are not significant. Type_of_farm
is not significant in any of the interactions or as a main effect. Year is only
significant as a main effect. The two-way interaction of Treatment and
Simplified_Boundary is significant. The other two-way interactions are not
significant.

# Model 1: LMM with 2-way interaction (T x S_B) and main-effect (Y)

## Description

We only include the significant terms, the 2-way interaction of Treatment and Simplified_Boundary, and the main-effect of Year.

## Diagnostics and ANOVA table

```{r, fig.width=9, fig.height=4}
# AIC(mod.lm); BIC(mod.lm)
mod.lm <- lmer(Mean_seeds_viable ~ Year + Treatment*Simplified_Boundary + (1|Farm), data=pc_mod)
# Has 5 outliers
# AIC(mod.lm); BIC(mod.lm) # Mixed result but BIC down more than AIC is up so all good.
# residplot(mod.lm, newwd = FALSE)
# summary(mod.lm)
simres <- simulateResiduals(fittedModel = mod.lm)
plot(simres) 
# rm(simres)
kable(Anova(mod.lm),digits=6, format = "pandoc")
```

There are 5 outliers (ignored for now).

## Model predictions

```{r}
emmip(mod.lm, Simplified_Boundary ~ Treatment | Year, 
      tlab=" ",
      xlab="", 
      ylab="Predicted mean number\nof viable seeds", 
      type = "response", CIs = TRUE, adjust="fdr") + theme_bw()

# emm <- emmeans(mod.lm, ~Simplified_Boundary|Treatment:Year, adjust="fdr") 
# emm %>% as.data.frame() %>% 
#   dplyr::select(-SE, -df) %>%
#   rename(`Predicted mean`=emmean) %>%
#   kable(format="pandoc", digits=1)
# 
# contrast(emm, method="pairwise", adjust="fdr", type = "response") %>%
#   as.data.frame() %>% 
#   dplyr::select(-SE, -df) %>%
#   kable(format="pandoc", digits=3)

emmip(mod.lm, Simplified_Boundary ~ Treatment,
      tlab=" ",
      xlab="", 
      ylab="Predicted mean number\nof viable seeds", 
      type = "response", CIs = TRUE, adjust="fdr") + theme_bw()

emm <- emmeans(mod.lm, ~Simplified_Boundary|Treatment, adjust="fdr") 
contrast(emm, method="pairwise", adjust="fdr", type = "response") %>%
  as.data.frame() %>% 
  dplyr::select(-SE, -df) %>%
  kable(format="pandoc", digits=3)

emm <- emmeans(mod.lm, ~Treatment|Simplified_Boundary, adjust="fdr") 
contrast(emm, method="pairwise", adjust="fdr", type = "response") %>%
  as.data.frame() %>% 
  dplyr::select(-SE, -df) %>%
  kable(format="pandoc", digits=3)

emmip(mod.lm, ~Year,
      xlab="", 
      ylab="Predicted mean number\nof viable seeds", 
      type = "response", CIs = TRUE, adjust = "fdr") + theme_bw()

emm <- emmeans(mod.lm, ~Year, adjust="fdr") 
contrast(emm, method="pairwise", adjust="fdr", type = "response") %>%
  as.data.frame() %>% 
  dplyr::select(-SE, -df) %>%
  kable(format="pandoc", digits=3)

# emm %>% as.data.frame() %>% 
#   dplyr::select(-SE, -df) %>%
#   rename(`Predicted mean`=emmean) %>%
#   kable(format="pandoc", digits=1)

```

## Graph of data with model predictions

```{r, fig.height=4.5, fig.width=6}
preddata <- pc_mod %>% dplyr::select(
  Treatment,
  Simplified_Boundary,
  Year,
  Farm,
  Mean_seeds_viable
)

emm <- emmeans(mod.lm, ~Simplified_Boundary:Treatment|Year, adjust="fdr") 
emm_ci <- emm %>% as.data.frame() %>% 
  rename(Predicted_mean=emmean) %>%
  dplyr::select(-SE, -df) 

# emm_ci %>% kable(format="pandoc", digits=3)

set.seed(1989)
preddata %>% 
  ggplot(aes(x=Treatment, y=Mean_seeds_viable, color=Treatment:Simplified_Boundary)) + 
  geom_jitter(height=0, width=0.2, size=0.5, alpha=0.2) + 
  geom_point(aes(y = Predicted_mean, color=Treatment:Simplified_Boundary), size = 1.4, data = emm_ci) +
  geom_errorbar(aes(y=Predicted_mean, ymin = lower.CL, ymax = upper.CL), width = 0.2, linewidth= 0.4, data = emm_ci) + 
  facet_grid(Year~Simplified_Boundary) + theme_bw(base_size=11) +
  labs(x="", y="Mean number of viable seeds", color="") +
  scale_color_manual(values = c("chocolate",
                                '#3B528BFF',
                                "darkgreen",  
                                "#440154FF",
                                "#DABF4AFF",
                                "#21908CFF",
                                "#5DC863FF",
                                "darkred"
                                )) + 
  theme(legend.position = "none") +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(colour = "black")
  ) 
```

# Model 2: LMM with 2-way interaction (T x S_B) and main-effect (Y) and 5 outliers removed

## Description

The model is the same as model 1 except the five outliers and removed from the data. The outliers are:

```{r}
# AIC(mod.lm); BIC(mod.lm)
# mod.lm <- lmer(Mean_seeds_viable ~ Year + Treatment*Simplified_Boundary + (1|Farm), data=pc_mod)
# Has 5 outliers
# AIC(mod.lm); BIC(mod.lm) # Mixed result but BIC down more than AIC is up so all good.
# residplot(mod.lm, newwd = FALSE)
# summary(mod.lm)
simres <- simulateResiduals(fittedModel = mod.lm)
#  plot(simres)
# out <- testOutliers(simres)
# out
# plotResiduals(simres, form = 1:nrow(pc_mod))
# which(simres$scaledResiduals < 0.008 | simres$scaledResiduals > 0.992)/214
tmp <- pc_mod[which(simres$scaledResiduals < 0.008 | simres$scaledResiduals > 0.992), ]

tmp  %>% 
  dplyr::select(-Mean_seeds_viable_rounded) %>%
  kable(format="pandoc", digits=1)
rm(tmp)

pc_mod_sub <- pc_mod[-which(simres$scaledResiduals < 0.008 | simres$scaledResiduals > 0.992), ]
```

## Diagnostics and ANOVA table

```{r, fig.width=9, fig.height=4}
mod.lm <- lmer(Mean_seeds_viable ~ Year + Treatment*Simplified_Boundary + (1|Farm), data=pc_mod_sub)
# residplot(mod.lm, newwd = FALSE)
# summary(mod.lm)
 simres <- simulateResiduals(fittedModel = mod.lm)
 plot(simres)
 
# rm(simres)
kable(Anova(mod.lm),digits=6, format = "pandoc")
```

## Model prediction

```{r}
emmip(mod.lm, Simplified_Boundary ~ Treatment | Year, 
      tlab=" ",
      xlab="", 
      ylab="Predicted mean number\nof viable seeds", 
      type = "response", CIs = TRUE, adjust="fdr") + theme_bw()

emm <- emmeans(mod.lm, ~Simplified_Boundary|Treatment:Year, adjust="fdr") 
# emm %>% as.data.frame() %>% 
#   dplyr::select(-SE, -df) %>%
#   rename(`Predicted mean`=emmean) %>%
#   kable(format="pandoc", digits=1)

# contrast(emm, method="pairwise", adjust="fdr", type = "response") %>%
#   as.data.frame() %>% 
#   dplyr::select(-SE, -df) %>%
#   kable(format="pandoc", digits=3)

emmip(mod.lm, Simplified_Boundary ~ Treatment,
      tlab=" ",
      xlab="", 
      ylab="Predicted mean number\nof viable seeds", 
      type = "response", CIs = TRUE, adjust="fdr") + theme_bw()

emm <- emmeans(mod.lm, ~Simplified_Boundary|Treatment, adjust="fdr") 

contrast(emm, method="pairwise", adjust="fdr", type = "response") %>%
  as.data.frame() %>% 
  dplyr::select(-SE, -df) %>%
  kable(format="pandoc", digits=3)

emm <- emmeans(mod.lm, ~Treatment|Simplified_Boundary, adjust="fdr") 
contrast(emm, method="pairwise", adjust="fdr", type = "response") %>%
  as.data.frame() %>% 
  dplyr::select(-SE, -df) %>%
  kable(format="pandoc", digits=3)

emmip(mod.lm, ~Year,
      xlab="", 
      ylab="Predicted mean number\nof viable seeds", 
      type = "response", CIs = TRUE, adjust = "fdr") + theme_bw()

emm <- emmeans(mod.lm, ~Year, adjust="fdr") 

contrast(emm, method="pairwise", adjust="fdr", type = "response") %>%
  as.data.frame() %>% 
  dplyr::select(-SE, -df) %>%
  kable(format="pandoc", digits=3)
```

## Graph of data with model predictions

```{r, fig.height=4.5, fig.width=6}
preddata_sub <- pc_mod_sub %>% dplyr::select(
  Treatment,
  Simplified_Boundary,
  Year,
  Farm,
  Mean_seeds_viable
)

emm <- emmeans(mod.lm, ~Simplified_Boundary:Treatment|Year, adjust="fdr") 
emm_ci <- emm %>% as.data.frame() %>% 
  rename(Predicted_mean=emmean) %>%
  dplyr::select(-SE, -df) 

# emm_ci %>% kable(format="pandoc", digits=3)

set.seed(1989)
preddata_sub %>% 
  ggplot(aes(x=Treatment, y=Mean_seeds_viable, color=Treatment:Simplified_Boundary)) + 
  geom_jitter(height=0, width=0.2, size=0.5, alpha=0.2) + 
  geom_point(aes(y = Predicted_mean, color=Treatment:Simplified_Boundary), size = 1.4, data = emm_ci) +
  geom_errorbar(aes(y=Predicted_mean, ymin = lower.CL, ymax = upper.CL), width = 0.2, linewidth= 0.4, data = emm_ci) + 
  facet_grid(Year~Simplified_Boundary) + theme_bw(base_size=11) +
  labs(x="", y="Mean number of viable seeds", color="") +
  scale_color_manual(values = c("chocolate",
                                '#3B528BFF',
                                "darkgreen",  
                                "#440154FF",
                                # "#FDE725FF",
                                "#DABF4AFF",
                                "#21908CFF",
                                "#5DC863FF",
                                "darkred"
                                )) + 
  theme(legend.position = "none") +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(colour = "black")
  ) 
```

# Model 3: LMM with 3-way interaction (T x S_B x Y)

## Description

We go back to Model 0 and only remove Type_of_farm, so we have the 3-way interaction of Year, Treatment and Simplified_Boundary.

## Diagnostics and ANOVA table

```{r, fig.width=9, fig.height=4}
mod.lm <- lmer(Mean_seeds_viable ~ Treatment*Simplified_Boundary*Year + (1|Farm), data=pc_mod)
# AIC(mod.lm); BIC(mod.lm)
# residplot(mod.lm, newwd = FALSE)
# summary(mod.lm)
 simres <- simulateResiduals(fittedModel = mod.lm)
 plot(simres)
# rm(simres)
kable(Anova(mod.lm),digits=4, format = "pandoc")
```

## Model predictions

```{r}
emmip(mod.lm, Simplified_Boundary ~ Treatment | Year,
      tlab=" ",
      xlab="", 
      ylab="Predicted mean number\nof viable seeds", 
      type = "response", CIs = TRUE, adjust="fdr") + theme_bw()

emmip(mod.lm, Simplified_Boundary ~ Treatment,
      tlab=" ",
      xlab="", 
      ylab="Predicted mean number\nof viable seeds", 
      type = "response", CIs = TRUE, adjust="fdr") + theme_bw()

emm <- emmeans(mod.lm, ~Simplified_Boundary|Treatment, adjust="fdr") 

contrast(emm, method="pairwise", adjust="fdr", type = "response") %>%
  as.data.frame() %>% 
  dplyr::select(-SE, -df) %>%
  kable(format="pandoc", digits=3)

emm <- emmeans(mod.lm, ~Treatment|Simplified_Boundary, adjust="fdr") 
contrast(emm, method="pairwise", adjust="fdr", type = "response") %>%
  as.data.frame() %>% 
  dplyr::select(-SE, -df) %>%
  kable(format="pandoc", digits=3)

emmip(mod.lm, ~Year,
      xlab="", 
      ylab="Predicted mean number\nof viable seeds", 
      type = "response", CIs = TRUE, adjust = "fdr") + theme_bw()

emm <- emmeans(mod.lm, ~Year, adjust="fdr") 
contrast(emm, method="pairwise", adjust="fdr", type = "response") %>%
  as.data.frame() %>% 
  dplyr::select(-SE, -df) %>%
  kable(format="pandoc", digits=3)
```

## Graph of data with model predictions

```{r, fig.height=4.5, fig.width=6}
preddata <- pc_mod %>% dplyr::select(
  Treatment,
  Simplified_Boundary,
  Year,
  Farm,
  Mean_seeds_viable
)

emm <- emmeans(mod.lm, ~Simplified_Boundary:Treatment|Year, adjust="fdr") 
emm_ci <- emm %>% as.data.frame() %>% 
  rename(Predicted_mean=emmean) %>%
  dplyr::select(-SE, -df) 

set.seed(1989)
preddata %>% 
  ggplot(aes(x=Treatment, y=Mean_seeds_viable, color=Treatment:Simplified_Boundary)) + 
  geom_jitter(height=0, width=0.2, size=0.5, alpha=0.2) + 
  geom_point(aes(y = Predicted_mean, color=Treatment:Simplified_Boundary), size = 1.4, data = emm_ci) +
  geom_errorbar(aes(y=Predicted_mean, ymin = lower.CL, ymax = upper.CL), width = 0.2, linewidth= 0.4, data = emm_ci) + 
  facet_grid(Year~Simplified_Boundary) + theme_bw(base_size=11) +
  labs(x="", y="Mean number of viable seeds", color="") +
  scale_color_manual(values = c("chocolate",
                                '#3B528BFF',
                                "darkgreen",  
                                "#440154FF",
                                # "#FDE725FF",
                                "#DABF4AFF",
                                "#21908CFF",
                                "#5DC863FF",
                                "darkred"
                                )) + 
  theme(legend.position = "none") +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(colour = "black")
  ) 
```

```{r, eval=FALSE}
# Poisson generalised linear mixed-effects model (Poisson GLMM)

# mod.lm <- glmer(Mean_seeds_viable_rounded ~ Type_of_farm*Treatment*Simplified_Boundary + (1|Year/Farm), family = poisson(link="identity"), data=pc_mod)
# AIC(mod.lm); BIC(mod.lm)
# mod.lm <- glmer(Mean_seeds_viable_rounded ~ Type_of_farm*Treatment*Simplified_Boundary + (1|Year) + (1|Farm), family = poisson(link="identity"), data=pc_mod)
# AIC(mod.lm); BIC(mod.lm)
# mod.lm <- glmer(Mean_seeds_viable_rounded ~ Type_of_farm*Treatment*Simplified_Boundary + (1|Farm), family = poisson(link="identity"), data=pc_mod)
# AIC(mod.lm); BIC(mod.lm)
mod.lm <- glmer(Mean_seeds_viable_rounded ~ Type_of_farm*Treatment*Simplified_Boundary + (1|Year), family = poisson(link="identity"), data=pc_mod)
# AIC(mod.lm); BIC(mod.lm)
# mod.lm <- glmer(Mean_seeds_viable_rounded ~ Type_of_farm*Treatment*Simplified_Boundary + (1|Farm/Year), family = poisson(link="identity"), data=pc_mod)
# AIC(mod.lm); BIC(mod.lm)
# residplot(mod.lm, newwd = FALSE)
# summary(mod.lm)
simres <- simulateResiduals(fittedModel = mod.lm)
plot(simres) # Not good
rm(simres)
```

