---
title: "B4BI Pak Choi Seed Data Graphs and Models for paper"
author: "Heather Jenkins"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---
<!-- output: -->
<!--   word_document: -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=4.5, dpi=600, warning=FALSE, message=FALSE, echo=FALSE)
options(scipen=999)
library(lme4)
library(tidyverse)
library(readxl)
library(janitor)
library(dplyr)
library(car)
library(emmeans)
library(predictmeans)
library(stringr)
library(GGally)
library(gridExtra)
library(pander)
library(kableExtra)
panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'left', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)
options(knitr.kable.NA = '')
```


# Overview

This document contains some analysis of the previously cleaned seed data.

This analysis was performed using  `r version$version.string`. The *tidyverse*
suite of packages was used for data manipulation and visualisation. The
*emmeans* package was used for model predictions.

# Data Loading

+ The cleaned and summarised (over the unbalanced reps) data is read in   
+ Some of the factor levels are renamed for the paper (can do this differently)  

<br>

```{r}
load(file="Cleaned_datasets/pc_seed_data_summarised_cleaned.rda")
  
pc_sum <- pc_sum %>% mutate(Treatment=fct_recode(Treatment, "HP"="Hand pollinated"),
                            Simplified_Boundary=fct_recode(Simplified_Boundary, 
                                        "Old native SNH"="Old native",
                                        "New native SNH"="New native",
                                        "On-farm bare fence"="Bare fence",
                                        "Off-farm bare fence"="Control farm"
                                        ))

str(pc_sum)
```

We next remove the data with 'one year native' and 'two year native' in 2023-2024 removed. So 'New native' only has

+ One year old in 2021-2022  
+ Two year old in 2022-2023  
+ Three year old in 2023-2024  

Note that is data is all from Arable farms (Garrickfield and Limbrick), and there are only 4 observations removed.

Tables of counts before and after the observations are removed are below:

```{r}
# str(pc_mod)
table(pc_sum$Year, pc_sum$Native_new_age)
pc_sum <- pc_sum %>% filter( !( (Native_new_age %in% c("One-year old native", "Two-year old native")) & Year=="2023-2024" ) )
table(pc_sum$Year, pc_sum$Native_new_age)
```

# Graphs

## Dairy or Arable added to the graphs

First we look at the data in a single graph and then with the farm types separated.

```{r, fig.height=4.75}
set.seed(1989)

pc_sum %>% 
  ggplot(aes(x=Treatment, y=Mean_seeds_viable, color=Treatment:Type_of_farm)) + 
  geom_jitter(height=0, width=0.3, size=1) + 
  # geom_point(aes(y = mean_viable), size=2, color=1, data = pc_stats) +
  facet_grid(Year~Simplified_Boundary) + theme_bw(base_size=11) +
  labs(x="", y="Mean number of viable seeds", color="") +
  theme(axis.text.x = element_text(angle=20, hjust=1.0, vjust=1.0)) +  
  scale_color_manual(values = c("chocolate",
                                "#21908CFF",
                                "#FDE725FF",
                                '#3B528BFF',
                                "#5DC863FF",
                                "#440154FF")) + 
  theme(
  legend.position = "bottom",
  legend.margin = margin(t = -10),
  legend.spacing.y = unit(0.1, "cm"),
  legend.key.height = unit(0.3, "cm")
) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(colour = "black")
  )
```

<br>

```{r}
set.seed(1989)
pc_sum %>% 
  filter(Type_of_farm=="Arable") %>% droplevels() %>%
  ggplot(aes(x=Treatment, y=Mean_seeds_viable, color=Treatment)) + 
  geom_jitter(height=0, width=0.3, size=1) + 
  facet_grid(Year~Simplified_Boundary) + theme_minimal() +
  labs(x="", y="Mean number of viable seeds",  title="Arable farms") +
  theme(axis.text.x = element_text(angle=20, hjust=1.0, vjust=1.0)) + 
  theme(legend.position = "none")

set.seed(1989)
pc_sum %>% 
  filter(Type_of_farm!="Arable") %>% droplevels() %>%
  ggplot(aes(x=Treatment, y=Mean_seeds_viable, color=Treatment)) + 
  geom_jitter(height=0, width=0.3, size=1) + 
  facet_grid(Year~Simplified_Boundary) + theme_minimal() +
  labs(x="", y="Mean number of viable seeds", title="Dairy farms") +
  theme(axis.text.x = element_text(angle=20, hjust=1.0, vjust=1.0)) + 
  theme(legend.position = "none")

# with(pc_sum, table(Farm, Simplified_Boundary)) # Sent to Brad to ask about Off-farm
```

# Modelling

## Comments

+ We remove the Control farms and the Closed treatment so we can see if there is a difference between HP and Open.  
+ Fitting a zero-inflated model with the Closed treatment data included did not converge. Since this was not the "main comparison of interest" I did not spend a lot time trying to make it work.  
+ The data is separated into Arable and Dairy for modelling purposes.   
+ Again I ran into problems with trying to fit a single model.  

## Model descriptions

We fit linear mixed-effect models to the two sets of data. The response is
Mean_seeds_viable. The fixed-effects are Treatment and Simplified_Boundary plus
their two-way interaction. Year is not really of interest (I hope?) so we make
it a random effect. There is also a random effect for Farm nested in Year. This
means that there are 3 levels to the Year random effect, and 24 for the
Farm:Year random effect (and it doesn't matter that we don't have data from
every farm for each year).

## Graphs of data used in models


```{r, fig.height=4.25}
pc_mod <- pc_sum %>% 
  filter(Simplified_Boundary != "Off-farm bare fence") %>% 
  filter(Treatment != "Closed") %>% 
  droplevels()

pc_mod_a <- pc_mod %>% 
  filter(Type_of_farm == "Arable") %>% 
  droplevels()

pc_mod_d <- pc_mod %>% 
  filter(Type_of_farm != "Arable") %>% 
  droplevels()

set.seed(1989)
pc_mod_a %>% 
  ggplot(aes(x=Treatment, y=Mean_seeds_viable, color=Treatment)) + 
  geom_jitter(height=0, width=0.3, size=1) + 
  facet_grid(Year~Simplified_Boundary) + theme_bw() +
  ylim(c(0,23.1)) +
  labs(x="", y="Mean number of viable seeds", color="", title="Arable farms, data used in model") +
  theme(axis.text.x = element_text(angle=20, hjust=1.0, vjust=1.0)) +  
  scale_color_manual(values=c("darkred", "#CD9600")) +
  theme(legend.position = "none")

set.seed(1989)
pc_mod_d %>% 
  ggplot(aes(x=Treatment, y=Mean_seeds_viable, color=Treatment)) + 
  geom_jitter(height=0, width=0.3, size=1) + 
  facet_grid(Year~Simplified_Boundary) + theme_bw() +
  ylim(c(0,23.1)) +
  labs(x="", y="Mean number of viable seeds", color="", title="Dairy farms, data used in model") +
  theme(axis.text.x = element_text(angle=20, hjust=1.0, vjust=1.0)) +  
  scale_color_manual(values=c("#F8766D", "#00BA38")) +
  theme(legend.position = "none")
```

## Results Arable farms

```{r, eval=FALSE}
# Results the same if this data used instead
pc_mod_a_sub <- pc_mod_a %>% 
  filter(Mean_seeds_viable > 5) %>% 
  filter(!(Mean_seeds_viable==9.5 & Treatment=="Hand pollinated")) 
```

The two outliers are not removed from the data. Removing them better satisfies
modelling assumptions, but does not change the story (results not shown).

The ANOVA table is:

<br>

```{r}
mod.lm <- lmer(Mean_seeds_viable ~ Treatment*Simplified_Boundary + (1|Year/Farm), data=pc_mod_a)
# mod.lm <- lmer(Mean_seeds_viable ~ Treatment*Simplified_Boundary + (1|Year/Farm), data=pc_mod_a_sub)
# mod.lm <- lmer(Mean_seeds_viable ~ Treatment*Simplified_Boundary + (1|Farm/Year), data=pc_mod_a_sub)
# mod.lm <- lmer(Mean_seeds_viable ~ Treatment*Simplified_Boundary + (1|Farm)+(1|Year), data=pc_mod_a_sub)

# summary(mod.lm)
kable(Anova(mod.lm),digits=6, format = "pandoc")
# AIC(mod.lm); BIC(mod.lm)
#residplot(mod.lm, newwd = FALSE)
```

<br>

The two-way interaction is significant. The modelled/predicted means and 95% confidence intervals are graphed and given
in a table. Note that the post-hoc pairwise comparison p-values were adjusted
using the FDR correction to account for multiple tests.

<br>

```{r, fig.height=2.5}
emmip(mod.lm, Simplified_Boundary ~ Treatment, 
      tlab=" ",
      xlab="", 
      ylab="Predicted mean number\nof viable seeds", 
      type = "response", CIs = TRUE) + theme_bw()

# emmip(mod.lm, Treatment ~ Simplified_Boundary, 
#       tlab="",
#       xlab="", 
#       ylab="Predicted mean number of viable seeds", 
#       type = "response", CIs = TRUE) + theme_bw()

emm <- emmeans(mod.lm, ~Simplified_Boundary|Treatment, adjust="fdr") 
emm %>% as.data.frame() %>% 
  dplyr::select(-SE, -df) %>%
  rename(`Predicted mean`=emmean) %>%
  kable(format="pandoc", digits=1)
```

<br>

```{r, fig.height=3.75}
plot(emm) + labs(x="Predicted mean number of viable seeds", y="Simplified boundary")
contrast(emm, method="pairwise", adjust="fdr", type = "response")
# mod.pm <- predictmeans(mod.lm, "Treatment:Simplified_Boundary", adj="fdr", pairwise=TRUE, newwd = FALSE) 
# mod.pm$mean_table
```

<br>

The table above shows pairwise Simplified_Boundary comparisons for each of the
Treatments. There is no difference for Hand pollinated.
For Open, the 'New native SNH' and 'Old native SNH' each have approximately 4 more viable
seeds on average than the Bare fence. There is no difference between 'New native
SNH' and 'Old native SNH'.

## Results Dairy farms

The ANOVA table is:

<br>

```{r}
mod.lm <- lmer(Mean_seeds_viable ~ Simplified_Boundary*Treatment + (1|Year/Farm), data=pc_mod_d)

# summary(mod.lm)
kable(Anova(mod.lm),digits=3, format = "pandoc")
# AIC(mod.lm); BIC(mod.lm)
# residplot(mod.lm, newwd = FALSE)
```

<br>

The two-way interaction is significant. The modelled/predicted means and 95% confidence intervals are graphed and given
in a table.

<br>

```{r, fig.height=2.5}
emmip(mod.lm, Simplified_Boundary ~ Treatment, 
      tlab=" ",
      xlab="", 
      ylab="Predicted mean number\nof viable seeds", 
      type = "response", CIs = TRUE) + theme_bw()

emm <- emmeans(mod.lm, ~Simplified_Boundary|Treatment, adjust="fdr") 
emm %>% as.data.frame() %>% 
  dplyr::select(-SE, -df) %>%
  rename(`Predicted mean`=emmean) %>%
  kable(format="pandoc", digits=1)
```

<br>

```{r, fig.height=3.75}
plot(emm) + labs(x="Predicted mean number of viable seeds", y="Simplified boundary")
contrast(emm, method="pairwise", adjust="fdr", type = "response")
# mod.pm <- predictmeans(mod.lm, "Treatment:Simplified_Boundary", adj="fdr", pairwise=TRUE, newwd = FALSE) 
# mod.pm$mean_table
```

<br>

The table above shows pairwise Simplified_Boundary comparisons for each of the Treatments. There is no difference for Hand pollinated.

For Open, the 'New native SNH' have approximately 3 more viable seeds on average than
the Bare fence. The 'Old native SNH' have approximately 4 more viable seeds on average
than the Bare fence. There is no difference between 'New native SNH' and 'Old native SNH'.
