---
title: "Final pakchoi diversity analyses"
author: "Eddy Dowle"
date: "2025-09-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lme4)
library(readxl)
library(svglite)
library(vegan)
library(iNEXT.3D)
library(ggpubr)
library(lmerTest)
library(emmeans)
library(RLRsim)
library(glmmTMB)
library(BiodiversityR)
library(ggpmisc)
library(pairwiseAdonis)
library(ggrepel)
library(ggvegan)
library(ggforce)
```

## Pak choi 

Diversity analyses, and summary plots

```{r cars}
#corrected dataset with raw counts
data<-read_excel('C:/Users/hrlexd/Dropbox/PlantAndFood (1)/B4BI/Pakchoi/Copy of Pak choi pollinators raw visits Brad correctedupdatedEddy.xlsx',sheet='Pak choi raw visits')

#filter out things that are 2018 in Boundary age column
data<-data %>% filter(`Boundary age`!=2018)


brads_col_2<-read_excel('C:/Users/hrlexd/Dropbox/PlantAndFood (1)/B4BI/Pakchoi/Eddy Pie chart short.xlsx',sheet='Colours')
#tidy up some names etc
data$Boundary_descriptor<-gsub('one year old','One year old',data$Boundary_descriptor)
data$Boundary_descriptor<-gsub('two year old','Two year old',data$Boundary_descriptor)
data$Boundary_descriptor<-gsub('three year old','Three year old',data$Boundary_descriptor)
data$Boundary_descriptor<-gsub('Bare_fence','Bare fence',data$Boundary_descriptor)
data$Boundary_descriptor<-gsub('Control_farm','Control farm',data$Boundary_descriptor)

#summarising counts to clean up dataset etc for box plots
data_box<-data %>% select(-`Broad Insect category`,-Origin,-`Boundary age`,-`Sum of Insect_Count`,-`Flower count`) %>% group_by(`Farm type`,Farm_name,Season,Boundary_descriptor,`Insect key number`,`New Inseck key name`) %>% summarise(`Sum of Insects/100 flowers`=sum(`Sum of Insects/100 flowers`) )

#counts per for below graph of abundance estimate per site
data_per<-data_box%>% mutate(Boundary_year=paste0(Boundary_descriptor,'_',Season)) %>% select(-`Insect key number`,-Boundary_descriptor,-Season) %>% group_by(`Farm type`,Farm_name,Boundary_year,`New Inseck key name`) %>% summarise(total_count=sum(`Sum of Insects/100 flowers`)) %>% ungroup() %>% group_by(`Farm type`,Farm_name,Boundary_year) %>% mutate(per=100*total_count/sum(total_count)) %>% ungroup()

data_per %>%   ggplot(aes(x = Boundary_year, y = per, fill = `New Inseck key name`)) +
  geom_bar(position='stack',stat='identity',width=0.4) +
  facet_grid(`Farm type`~Farm_name, scales = "free", space = "free") +
  theme_bw(base_size = 12) +
  theme(strip.text.x = element_text(angle = 90,size=7),axis.text.x = element_text(angle = 90,size=7,hjust=1))+
  ggtitle(paste('test'))+
 theme(legend.position="none")


mydf<- data_box %>% mutate(Boundary_year=paste0(Boundary_descriptor,'_',Season))

#setting up for brads colour scheme
#relevel to brads order
brads_col_2$`New Inseck key name` <- factor(brads_col_2$`New Inseck key name`, levels=brads_col_2$`New Inseck key name`[order(brads_col_2$`Insect key number`)], ordered=TRUE)

#relevel to brads order
mydf$`New Inseck key name` <- forcats::fct_relevel(mydf$`New Inseck key name` ,levels(brads_col_2$`New Inseck key name`))

```

## Diversity statistics

Bar charts to view diversity over different groups

```{r pressure, echo=T}
##bar chart raw abundances .....
#cant just be raw count
#needs to be average by sampling number

mydf_byboundary_farm_count<-mydf %>% select(-`Insect key number`,-`Farm type`,-Boundary_year) %>% mutate(unique_ob=paste(Boundary_descriptor,Farm_name,Season)) %>% select(-Farm_name,-Season) %>% group_by(Boundary_descriptor,`New Inseck key name`) %>% summarise(count_farms_perobs=sum(`Sum of Insects/100 flowers`)/n_distinct(unique_ob)) 
mydf_byboundary_farm_count$Boundary_descriptor <- factor(mydf_byboundary_farm_count$Boundary_descriptor,levels=c('Control farm','Bare fence','One year old','Two year old','Three year old','Old'),ordered = T)

mydf_byboundary_farm_count %>%   ggplot(aes(x = Boundary_descriptor, y = count_farms_perobs, fill = `New Inseck key name`)) +
  geom_bar(position='stack',stat='identity',width=0.4) +
  # facet_grid(`Farm type`~Farm_name, scales = "free", space = "free") +
  theme_bw(base_size = 12) +
  theme(strip.text.x = element_text(angle = 90,size=7),axis.text.x = element_text(angle = 90,size=7,hjust=1))+
  scale_fill_manual(values=with(brads_col_2,setNames(Brad_colours,`New Inseck key name`)))+
  ggtitle(paste('Counts by boundary'))+ylab('Average count per 100 flowers')+xlab('Boundary type')+
  labs(fill = "Insect  group")+
  theme(
    legend.text = element_text(size = 6), 
    legend.title = element_text(size = 6),
    legend.key.size = unit(1,"line"))  

#bar chart relative abundances % bare fence, control year 1, year 2 year3 and old
mydf_byboundary<-mydf%>% select(-`Insect key number`,-Season,-`Farm type`,-Farm_name,-Boundary_year) %>% group_by(Boundary_descriptor,`New Inseck key name`) %>% summarise(total_count=sum(`Sum of Insects/100 flowers`)) %>% ungroup() %>% group_by(Boundary_descriptor) %>% mutate(per=100*total_count/sum(total_count)) %>% ungroup()

#recolour and tidy
mydf_byboundary$Boundary_descriptor <- factor(mydf_byboundary$Boundary_descriptor,levels=c('Control farm','Bare fence','One year old','Two year old','Three year old','Old'),ordered = T)

mydf_byboundary %>%   ggplot(aes(x = Boundary_descriptor, y = per, fill = `New Inseck key name`)) +
  geom_bar(position='stack',stat='identity',width=0.4) +
  # facet_grid(`Farm type`~Farm_name, scales = "free", space = "free") +
  theme_bw(base_size = 12) +
  theme(strip.text.x = element_text(angle = 90,size=7),axis.text.x = element_text(angle = 90,size=7,hjust=1),plot.title = element_text(size = 13))+
  scale_fill_manual(values=with(brads_col_2,setNames(Brad_colours,`New Inseck key name`)))+
  ggtitle(paste('Relative abundance by boundary'))+ylab('Proportion')+xlab('Boundary type')+
  labs(fill = "Insect  group")+
  theme(
    legend.text = element_text(size = 6), 
    legend.title = element_text(size = 6),
    legend.key.size = unit(1,"line"))  

```

## Bar charts split by season

Alpha diversity (below) suggests there might be a role of season, so just splitting that out here. Season makes sense as the second year had a lot less counts than the first. 

```{r bar_season, echo=T}
#Adding in season
##bar chart raw abundances .....
#cant just be raw count
#needs to be average by sampling number

mydf_byboundary_farm_count_yr<-mydf %>% select(-`Insect key number`,-`Farm type`,-Boundary_descriptor,-Season) %>% mutate(unique_ob=paste(Boundary_year,Farm_name)) %>% select(-Farm_name) %>% group_by(Boundary_year,`New Inseck key name`) %>% summarise(count_farms_perobs=sum(`Sum of Insects/100 flowers`)/n_distinct(unique_ob)) 

mydf_byboundary_farm_count_yr<-mydf_byboundary_farm_count_yr %>% separate_wider_delim(Boundary_year,'_',names=c("Boundary_descriptor","Season"))

mydf_byboundary_farm_count_yr$Boundary_descriptor <- factor(mydf_byboundary_farm_count_yr$Boundary_descriptor,levels=c('Control farm','Bare fence','One year old','Two year old','Three year old','Old'),ordered = T)

mydf_byboundary_farm_count_yr %>%   ggplot(aes(x = Boundary_descriptor, y = count_farms_perobs, fill = `New Inseck key name`)) +
  geom_bar(position='stack',stat='identity',width=0.4) +
  facet_grid(~Season, scales = "free") +
  theme_bw(base_size = 12) +
  theme(strip.text.x = element_text(angle = 0,size=7),axis.text.x = element_text(angle = 90,size=7,hjust=1),plot.title = element_text(size = 13))+
  scale_fill_manual(values=with(brads_col_2,setNames(Brad_colours,`New Inseck key name`)))+
  ggtitle(paste('Counts by boundary'))+ylab('Average count per 100 flowers')+xlab('Boundary type')+
  labs(fill = "Insect  group")+
  theme(
    legend.text = element_text(size = 6), 
    legend.title = element_text(size = 6),
    legend.key.size = unit(1,"line"))  

#adding in season 
#bar chart relative abundances % bare fence, control year 1, year 2 year3 and old

mydf_byboundary_yr<-mydf%>% select(-`Insect key number`,-Season,-`Farm type`,-Farm_name,-Boundary_descriptor) %>% group_by(Boundary_year,`New Inseck key name`) %>% summarise(total_count=sum(`Sum of Insects/100 flowers`)) %>% ungroup() %>% group_by(Boundary_year) %>% mutate(per=100*total_count/sum(total_count)) %>% ungroup()

mydf_byboundary_yr<-mydf_byboundary_yr %>% separate_wider_delim(Boundary_year,'_',names=c("Boundary_descriptor","Season"))

#recolour and tidy
mydf_byboundary_yr$Boundary_descriptor <- factor(mydf_byboundary_yr$Boundary_descriptor,levels=c('Control farm','Bare fence','One year old','Two year old','Three year old','Old'),ordered = T)

mydf_byboundary_yr %>%   ggplot(aes(x = Boundary_descriptor, y = per, fill = `New Inseck key name`)) +
  geom_bar(position='stack',stat='identity',width=0.4) +
  facet_wrap(~Season,scales="free") +
  theme_bw(base_size = 12) +
  theme(strip.text.x = element_text(angle = 0,size=7),axis.text.x = element_text(angle = 90,size=7,hjust=1),plot.title = element_text(size = 13))+
  scale_fill_manual(values=with(brads_col_2,setNames(Brad_colours,`New Inseck key name`)))+
  ggtitle(paste('Relative abundance by boundary'))+ylab('Proportion')+xlab('Boundary type')+
  labs(fill = "Insect  group")+
  theme(
    legend.text = element_text(size = 6), 
    legend.title = element_text(size = 6),
    legend.key.size = unit(1,"line"))  

```

## Just a base tests to show that there is significantly more counts by planting age

Just mucking around trying to think of ways to highlight the great raw counts in old, 3 years etc

```{r raw_counts, echo=T,warning = FALSE}
#switching to raw counts
data<-read_excel('C:/Users/hrlexd/Dropbox/PlantAndFood (1)/B4BI/Pakchoi/Copy of Pak choi pollinators raw visits Brad correctedupdatedEddy.xlsx',sheet='Pak choi raw visits')

#filter out things that are 2018 in Boundary age column
data<-data %>% filter(`Boundary age`!=2018)

brads_col_2<-read_excel('C:/Users/hrlexd/Dropbox/PlantAndFood (1)/B4BI/Pakchoi/Eddy Pie chart short.xlsx',sheet='Colours')
#tidy up some names etc
data$Boundary_descriptor<-gsub('one year old','One year old',data$Boundary_descriptor)
data$Boundary_descriptor<-gsub('two year old','Two year old',data$Boundary_descriptor)
data$Boundary_descriptor<-gsub('three year old','Three year old',data$Boundary_descriptor)
data$Boundary_descriptor<-gsub('Bare_fence','Bare fence',data$Boundary_descriptor)
data$Boundary_descriptor<-gsub('Control_farm','Control farm',data$Boundary_descriptor)

unique(data$Boundary_descriptor)
colnames(data)

#summarising counts to clean up dataset etc
mydf<-data %>% select(-`Broad Insect category`,-Origin,-`Boundary age`,-`Sum of Insects/100 flowers`) 

mydf<-mydf %>% mutate(unique_sample=paste(Boundary_descriptor,Farm_name,Season))

counts<-mydf %>% ungroup() %>% select(unique_sample,`New Inseck key name`,`Sum of Insect_Count`)

#transform wide
counts_wide<-counts %>% pivot_wider(names_from=`New Inseck key name`,values_from = `Sum of Insect_Count`)
counts_wide<-counts_wide %>% remove_rownames %>% column_to_rownames(var="unique_sample")

#meta data
meta_data<-read_excel('C:/Users/hrlexd/Dropbox/PlantAndFood (1)/B4BI/Pakchoi/Copy of Meta_data_pakchoiBH.xlsx',sheet='Meta_data_pakchoi')
rownames(meta_data) <- meta_data$unique_sample
meta_data<-meta_data %>% mutate(month_split=paste(Collection_month,Beg_Mid_end,sep='_'))

#brad colour scheme (control to old):
col_bound<-c('chocolate','#FDE725FF',"#5DC863FF","#21908CFF",'#3B528BFF',"#440154FF")
#col_bound

#ready to analyse!

insect_counts<-rowSums(counts_wide)

insect_counts_df<-data.frame(insect_counts)
insect_counts_df <- rownames_to_column(insect_counts_df, var = "unique_sample")
insect_counts_df<-left_join(insect_counts_df,meta_data)
insect_counts_df$Boundary_descriptor <- factor(insect_counts_df$Boundary_descriptor,levels=c('Control farm','Bare fence','One year old','Two year old','Three year old','Old'),ordered = T)
#split by season
plot_sppr <- ggplot(insect_counts_df, aes(x = Boundary_descriptor, y = insect_counts, fill = Boundary_descriptor)) +
  geom_boxplot() +
#  facet_wrap(~Season)+
  theme_bw()+
  scale_fill_manual(values=col_bound)+
  labs(x = "Boundary type",
       y = "Raw counts",
       title = "Raw counts by boundary",
       fill = "Boundary type")+
  theme(axis.text.x = element_text(angle = 90,size=7,hjust=1))

plot_sppr


#split by season
plot_sppr <- ggplot(insect_counts_df, aes(x = Boundary_descriptor, y = insect_counts, fill = Boundary_descriptor)) +
  geom_boxplot() +
  facet_wrap(~Season,scales="free")+
  theme_bw()+
  scale_fill_manual(values=col_bound)+
  labs(x = "Boundary type",
       y = "Raw counts",
       title = "Raw counts by boundary",
       fill = "Boundary type")+
  theme(axis.text.x = element_text(angle = 90,size=7,hjust=1))

plot_sppr

#issue with model convergence with some random effects so:
m0<-lmer(insect_counts ~ Boundary_descriptor + offset(log(`Flower count`)) +(1|Season) , data = meta_data)
m1<-lmer(insect_counts ~ Boundary_descriptor + offset(log(`Flower count`)) +(1|Season) +(1|Farm_name), data = meta_data)
#im getting a boundary (singular) fit warning with month_split so dropping for this

AIC(m0,m1)
anova(m0,m1)

#final model
m0<-lmer(insect_counts ~ Boundary_descriptor + offset(log(`Flower count`)) +(1|Season) +(1|Farm_name), data = meta_data)

emm2 <- emmeans(m0, ~ Boundary_descriptor,type = "response")
pairs(emm2,simple='each',  adjust = "tukey")

#running each as I think all comparisons is a bit over the top and probably repeating things
cld_rich <- multcomp::cld(emm2, Letters = letters, adjust = "tukey")
colnames(cld_rich)

cld_rich$Boundary_descriptor <- factor(cld_rich$Boundary_descriptor,levels=c('Control farm','Bare fence','One year old','Two year old','Three year old','Old'),ordered = T)

ggplot(cld_rich, aes(x = Boundary_descriptor, y = emmean)) +
  theme_bw()+
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  geom_point(size = 3,aes(colour=Boundary_descriptor)) +
  scale_colour_manual(values=col_bound)+
# not sure if groupings will make sense with facet wrap
#  geom_text(aes(label = .group, y = asymp.UCL + 0.2), vjust = 0, size = 5) +
  labs(y = "Estimated counts (adjusted for flowers)", x = "Boundary type", colour='Boundary type') +
  theme(axis.text.x = element_text(angle = 90,size=7,hjust=1))

#given how misrabley different they are we are probably safe with just an anova

count_aov <-aov(insect_counts ~ Boundary_descriptor + Season + offset(log(`Flower count`)) , data = meta_data)
summary(count_aov)
TukeyHSD(count_aov)

#cough cough bring 3 year down below 0.05 rather than like -.08-0.09 on the full model
```

## Alpha diversity

Species count using species count 
Im switching to raw counts here, one issue is that flower number varies so need to check the relationship between diversity and flower number

```{r raw_alpha, echo=T}
#Alpha diversity
#one matrix of the counts
#one metadata file

#species accumulation curve
speccurve<-specaccum(counts_wide,method='random',permutations=1000)
plot(speccurve)

#count species per property
sppr <- specnumber(counts_wide)
head(sppr)

#can plot that out
sppr_df <- sppr %>% 
  enframe() %>% 
  full_join(meta_data, by = c("name" = "unique_sample"))

sppr_df$Boundary_descriptor <- factor(sppr_df$Boundary_descriptor,levels=c('Control farm','Bare fence','One year old','Two year old','Three year old','Old'),ordered = T)

#just boundary
plot_sppr <- ggplot(sppr_df, aes(x = Boundary_descriptor, y = value, fill = Boundary_descriptor)) +
  geom_boxplot() +
  theme_bw()+
  scale_fill_manual(values=col_bound)+
  labs(x = "Boundary type",
       y = "Number of species per site",
       title = "Species richness: species count",
       fill = "Boundary type")
plot_sppr

#other significant factor was season
plot_sppr <- ggplot(sppr_df, aes(x = Boundary_descriptor, y = value, fill = Boundary_descriptor)) +
  geom_boxplot() +
facet_wrap(~Season)+
    scale_fill_manual(values=col_bound)+
  theme_bw()+
  labs(x = "Boundary type",
       y = "Number of species per site",
       title = "Species richness: species count",
       fill = "Boundary type")+
  theme(axis.text.x = element_text(angle = 90,size=7,hjust=1))

plot_sppr

#flower number:
plot_sppr <- ggplot(sppr_df, aes(x = `Flower count`, y = value)) +
  geom_point(aes(colour = Boundary_descriptor)) +
  scale_colour_manual(values=col_bound)+
  stat_poly_line() +
  stat_poly_eq() +
   theme_bw()+
  labs(x = "Flower Number",
       y = "Number of species per site",
       title = "Species richness: species count", colour = "Boundary type")+
  theme(axis.text.x = element_text(angle = 90,size=7,hjust=1))

plot_sppr

####season as a random effect
#date split by month (early, mid, late) and farm name
logLik(lmer(sppr ~ Boundary_descriptor + offset(log(`Flower count`))+(1|Season), data = meta_data))
m0<-lmer(sppr ~ Boundary_descriptor + offset(log(`Flower count`))+(1|Season), data = meta_data)
logLik(lmer(sppr ~ Boundary_descriptor + offset(log(`Flower count`))+(1|Season) + (1|Farm_name) , data = meta_data))
m1<-lmer(sppr ~ Boundary_descriptor + offset(log(`Flower count`))+(1|Season) + (1|Farm_name) , data = meta_data)
logLik(lmer(sppr ~ Boundary_descriptor + offset(log(`Flower count`))+(1|Season) + (1|Collection_month) + (1|Site), data = meta_data))
m2<-lmer(sppr ~ Boundary_descriptor + offset(log(`Flower count`))+(1|Season) + (1|Collection_month) + (1|Farm_name), data = meta_data)
logLik(lmer(sppr ~ Boundary_descriptor + offset(log(`Flower count`))+(1|Season) + (1|month_split) + (1|Farm_name), data = meta_data))
m3<-lmer(sppr ~ Boundary_descriptor + offset(log(`Flower count`))+(1|Season) + (1|month_split) + (1|Farm_name), data = meta_data)

AIC(m0,m1,m2,m3)
anova(m0,m1,m2,m3)

#this is all really marginal so be honest
#in my mind all these are so marginal but for explaining in a paper it may look better to say farm and month were included as random effects:

#so date and season as random effect
#month_split + farm name + season
m3<-lmer(sppr ~ Boundary_descriptor + offset(log(`Flower count`))+(1|Season) + (1|month_split) + (1|Farm_name), data = meta_data)
#back transform to species count
emm1 <- emmeans(m3, ~ Boundary_descriptor, type = "response") 
#pairwise test
pairs(emm1, simple='each', adjust = "tukey")
#pairs(emm1, by='Boundary_descriptor', adjust = "tukey")
#running each as I think all comparisons is a bit over the top and probably repeating things
cld_rich <- multcomp::cld(emm1, Letters = letters, adjust = "tukey")
colnames(cld_rich)

cld_rich$Boundary_descriptor <- factor(cld_rich$Boundary_descriptor,levels=c('Control farm','Bare fence','One year old','Two year old','Three year old','Old'),ordered = T)

ggplot(cld_rich, aes(x = Boundary_descriptor, y = emmean)) +
  theme_bw()+
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  geom_point(size = 3,aes(colour=Boundary_descriptor)) +
  scale_colour_manual(values=col_bound)+
# not sure if groupings will make sense with facet wrap
#  geom_text(aes(label = .group, y = asymp.UCL + 0.2), vjust = 0, size = 5) +
  labs(y = "Species richness (adjusted for flowers)", x = "Boundary type", colour='Boundary type') +
  theme(axis.text.x = element_text(angle = 90,size=7,hjust=1))

#but this is really a marginal on those random effects and if you just do a standard anova:
#flower count as a offset 
sppr_aov <-aov(sppr ~ Boundary_descriptor + Season + offset(log(`Flower count`)) , data = meta_data)
summary(sppr_aov)
TukeyHSD(sppr_aov)

#the significant pairs of boundary types are exactly the same between the full model and the anova
```

## Alpha diversity shannon

Alpha diversity using shannon

```{r shannon_alpha, echo=T}
shannondiv<-diversity(counts_wide)
head(shannondiv)


#can plot that out
shannondiv_df <- shannondiv %>% 
  enframe() %>% 
  full_join(meta_data, by = c("name" = "unique_sample"))

shannondiv_df$Boundary_descriptor <- factor(shannondiv_df$Boundary_descriptor,levels=c('Control farm','Bare fence','One year old','Two year old','Three year old','Old'),ordered = T)


#just boundary
plot_sppr <- ggplot(shannondiv_df, aes(x = Boundary_descriptor, y = value, fill = Boundary_descriptor)) +
  geom_boxplot() +
  theme_bw()+
  scale_fill_manual(values=col_bound)+
  labs(x = "Boundary type",
       y = "Number of species per site",
       title = "Species richness: shannon",
       fill = "Boundary type")
plot_sppr

#split by season
plot_sppr <- ggplot(shannondiv_df, aes(x = Boundary_descriptor, y = value, fill = Boundary_descriptor)) +
  geom_boxplot() +
  facet_wrap(~Season)+
  theme_bw()+
  scale_fill_manual(values=col_bound)+
  labs(x = "Boundary type",
       y = "Number of species per site",
       title = "Species richness: shannon",
       fill = "Boundary type")+
  theme(axis.text.x = element_text(angle = 90,size=7,hjust=1))

plot_sppr

####season as random effect
logLik(lmer(shannondiv ~ Boundary_descriptor+ offset(log(`Flower count`))+(1|Season), data = meta_data))
m0<-lmer(shannondiv ~ Boundary_descriptor+ offset(log(`Flower count`))+(1|Season), data = meta_data)
logLik(lmer(shannondiv ~ Boundary_descriptor+ offset(log(`Flower count`))+(1|Season) + (1|Farm_name) , data = meta_data))
m1<-lmer(shannondiv ~ Boundary_descriptor+ offset(log(`Flower count`))+(1|Season) + (1|Farm_name) , data = meta_data)
logLik(lmer(shannondiv ~ Boundary_descriptor+ offset(log(`Flower count`))+(1|Season) + (1|month_split) , data = meta_data))
m2<-lmer(shannondiv ~ Boundary_descriptor+ offset(log(`Flower count`))+(1|Season) + (1|month_split) , data = meta_data)
logLik(lmer(shannondiv ~ Boundary_descriptor + offset(log(`Flower count`))+(1|Season) + (1|month_split) + (1|Farm_name), data = meta_data))
m3<-lmer(shannondiv ~ Boundary_descriptor + offset(log(`Flower count`))+(1|Season) + (1|month_split) + (1|Farm_name), data = meta_data)

AIC(m0,m1,m2,m3) 
anova(m0,m1,m2,m3)
#m3 model which has date and site grouping....

m3<-lmer(shannondiv ~ Boundary_descriptor + offset(log(`Flower count`))+(1|Season) + (1|month_split) + (1|Farm_name), data = meta_data)

emm2 <- emmeans(m3, ~ Boundary_descriptor,type = "response")
pairs(emm2,simple='each',  adjust = "tukey")

#running each as I think all comparisons is a bit over the top and probably repeating things
cld_rich <- multcomp::cld(emm2, Letters = letters, adjust = "tukey")
colnames(cld_rich)

cld_rich$Boundary_descriptor <- factor(cld_rich$Boundary_descriptor,levels=c('Control farm','Bare fence','One year old','Two year old','Three year old','Old'),ordered = T)

ggplot(cld_rich, aes(x = Boundary_descriptor, y = emmean)) +
#  facet_wrap(~Season)+
  theme_bw()+
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  geom_point(size = 3,aes(colour=Boundary_descriptor)) +
  scale_colour_manual(values=col_bound)+
# not sure if groupings will make sense with facet wrap
#  geom_text(aes(label = .group, y = asymp.UCL + 0.2), vjust = 0, size = 5) +
  labs(y = "Estimated Shannon (adjusted for flowers)", x = "Boundary type", colour='Boundary type') +
  theme(axis.text.x = element_text(angle = 90,size=7,hjust=1))


#just runing an anova instead
#flower count as a offset 
sppr_aov <-aov(shannondiv ~ Boundary_descriptor + Season + offset(log(`Flower count`)) , data = meta_data)
summary(sppr_aov)
TukeyHSD(sppr_aov)


```

## Alpha diversity simpson

Alpha diversity using simpson

```{r simpson_alpha, echo=T}
simpsondiv<-diversity(counts_wide,index='simpson')
head(simpsondiv)

#can plot that out
simpsondiv_df <- simpsondiv %>% 
  enframe() %>% 
  full_join(meta_data, by = c("name" = "unique_sample"))

simpsondiv_df$Boundary_descriptor <- factor(simpsondiv_df$Boundary_descriptor,levels=c('Control farm','Bare fence','One year old','Two year old','Three year old','Old'),ordered = T)


#just boundary
plot_sppr <- ggplot(simpsondiv_df, aes(x = Boundary_descriptor, y = value, fill = Boundary_descriptor)) +
  geom_boxplot() +
  theme_bw()+
  scale_fill_manual(values=col_bound)+
  labs(x = "Boundary type",
       y = "Number of species per site",
       title = "Species richness: simpson",
       fill = "Boundary type")
plot_sppr

#split by season
plot_sppr <- ggplot(simpsondiv_df, aes(x = Boundary_descriptor, y = value, fill = Boundary_descriptor)) +
  geom_boxplot() +
  facet_wrap(~Season)+
  theme_bw()+
  scale_fill_manual(values=col_bound)+
  labs(x = "Boundary type",
       y = "Number of species per site",
       title = "Species richness: simpson",
       fill = "Boundary type")+
  theme(axis.text.x = element_text(angle = 90,size=7,hjust=1))

plot_sppr


####season as random effect
logLik(lmer(simpsondiv ~ Boundary_descriptor+ offset(log(`Flower count`))+(1|Season), data = meta_data))
m0<-lmer(simpsondiv  ~ Boundary_descriptor+ offset(log(`Flower count`))+(1|Season), data = meta_data)
logLik(lmer(simpsondiv ~ Boundary_descriptor+ offset(log(`Flower count`))+(1|Season) + (1|Farm_name) , data = meta_data))
m1<-lmer(simpsondiv ~ Boundary_descriptor+ offset(log(`Flower count`))+(1|Season) + (1|Farm_name) , data = meta_data)
logLik(lmer(simpsondiv ~ Boundary_descriptor+ offset(log(`Flower count`))+(1|Season) + (1|month_split) , data = meta_data))
m2<-lmer(simpsondiv ~ Boundary_descriptor+ offset(log(`Flower count`))+(1|Season) + (1|month_split) , data = meta_data)
logLik(lmer(simpsondiv ~ Boundary_descriptor+ offset(log(`Flower count`))+(1|Season) + (1|month_split) + (1|Farm_name), data = meta_data))
m3<-lmer(simpsondiv ~ Boundary_descriptor+ offset(log(`Flower count`))+(1|Season) + (1|month_split) + (1|Farm_name), data = meta_data)

AIC(m0,m1,m2,m3) #m3
anova(m0,m1,m2,m3)

m3<-lmer(simpsondiv ~ Boundary_descriptor+ offset(log(`Flower count`))+(1|Season) + (1|month_split) + (1|Farm_name), data = meta_data)

emm2 <- emmeans(m3, ~ Boundary_descriptor,type = "response")
pairs(emm2,simple='each',  adjust = "tukey")

#running each as I think all comparisons is a bit over the top and probably repeating things
cld_rich <- multcomp::cld(emm2, Letters = letters, adjust = "tukey")
colnames(cld_rich)

cld_rich$Boundary_descriptor <- factor(cld_rich$Boundary_descriptor,levels=c('Control farm','Bare fence','One year old','Two year old','Three year old','Old'),ordered = T)

ggplot(cld_rich, aes(x = Boundary_descriptor, y = emmean)) +
#  facet_wrap(~Season)+
  theme_bw()+
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  geom_point(size = 3,aes(colour=Boundary_descriptor)) +
  scale_colour_manual(values=col_bound)+
# not sure if groupings will make sense with facet wrap
#  geom_text(aes(label = .group, y = asymp.UCL + 0.2), vjust = 0, size = 5) +
  labs(y = "Estimated Shannon (adjusted for flowers)", x = "Boundary type", colour='Boundary type') +
  theme(axis.text.x = element_text(angle = 90,size=7,hjust=1))

#flower count as a offset 
sppr_aov <-aov(simpsondiv  ~ Boundary_descriptor + Season + offset(log(`Flower count`)) , data = meta_data)
summary(sppr_aov)
TukeyHSD(sppr_aov)

```

## Using renyi profiles
Tracking the difference between the two communities using a renyi profile method. Dont think we can use this in the paper as I cant get season, date or farm in there for the nesting. But thought it was interesting. 

The Renyi diversity profile at scale 0 reflects species richness, at scale 1 the Shannon index, at scale 2 the Simpson index and at scale Inf the Berger-Parker index (the dominance of the most abundant species):

If the profiles cross over then it may not be possible to compare diversity. I find this interesting as its the old samples that cross over. The 3 year olds don't. This follows the above shannon/simpson analysis that shows evenness is highest in the 3-year old samples. We cant actually include this in the final paper as its not taking into account flower number/farm/date sampled like the models above are. But I wanted to get your thoughts on it as I havent seen data like this. Though the fly one was similar when simpson was so high for apple as there was basically no diversity at apple so it was perfectly even. We should probably include the accumulation plot in the supplementary though (along with the innext plots).

Im thinking that this suggests a lot more rare species are coming into the plantings compared to the bare fence. So the evenness becomes quite lower compared to the bare fence/control plantings. Which feeds into that science paper we discussed the other day about bee species rather than bee number being important. 

```{r renyi_q1,echo=T}
meta_data$Boundary_descriptorf<-factor(meta_data$Boundary_descriptor, levels = c('Control farm','Bare fence',"One year old","Two year old","Three year old","Old"))
meta_data_bior<-as.data.frame(meta_data)
Renyi.1<-renyicomp(counts_wide,y=meta_data_bior,factor='Boundary_descriptorf', 
  scales=c(0, 0.25, 0.5, 1, 2, 4, 8, Inf), permutations=1000, plotit=F)
renyi.long1 <- renyicomp.long(Renyi.1, label.freq=1)
renyi.long1_dataframe<-data.frame(renyi.long1$labelit,renyi.long1$Scales,renyi.long1$Obs,renyi.long1$Diversity,renyi.long1$UPR,renyi.long1$LWR,renyi.long1$Grouping)
renyi.long1_dataframe$Boundary_descriptor<-factor(renyi.long1_dataframe$renyi.long1.Grouping, levels = c('Control farm','Bare fence',"One year old","Two year old","Three year old","Old"))

pl_renyi_across <- ggplot(data=renyi.long1_dataframe, aes(x=renyi.long1.Scales, y=renyi.long1.Diversity, ymax=renyi.long1.UPR, ymin=renyi.long1.LWR)) + 
    scale_x_discrete() +
    scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
    geom_line(data=renyi.long1_dataframe, 
              aes(x=renyi.long1.Obs, colour=Boundary_descriptor), 
              linewidth=2) +
    geom_point(data=subset(renyi.long1_dataframe, renyi.long1.labelit==TRUE), 
        aes(colour=Boundary_descriptor, shape=Boundary_descriptor), 
        size=5) +
    geom_ribbon(data=renyi.long1_dataframe, 
                aes(x=renyi.long1.Obs, colour=Boundary_descriptor, fill=after_scale(alpha(colour, 0.1))), 
                show.legend=FALSE) + 
     scale_colour_manual(values=col_bound)+
    theme_bw() +
    #scale_colour_npg() +
    labs(x=expression(alpha), y = "Diversity", colour = "Boundary type", shape = "Boundary type")
pl_renyi_across

#not great but that is conflated with the effect of season so break it up by season:
meta_data_bior_season<-meta_data_bior %>% filter(Season=='2023/24')
meta_data_bior_season %>% select(Boundary_descriptor) %>% group_by(Boundary_descriptor) %>% summarise(count = n())

counts_wide_filt <- counts_wide[rownames(counts_wide) %in% meta_data_bior_season$unique_sample, ]
counts_wide_filt <- counts_wide_filt[,colSums(counts_wide_filt) > 0]
Renyi.1<-renyicomp(counts_wide_filt,y=meta_data_bior_season,factor='Boundary_descriptorf', 
  scales=c(0, 0.25, 0.5, 1, 2, 4, 8, Inf), permutations=1000, plotit=F)
renyi.long1 <- renyicomp.long(Renyi.1, label.freq=1)
renyi.long1_dataframe<-data.frame(renyi.long1$labelit,renyi.long1$Scales,renyi.long1$Obs,renyi.long1$Diversity,renyi.long1$UPR,renyi.long1$LWR,renyi.long1$Grouping)
renyi.long1_dataframe$Boundary_descriptor<-factor(renyi.long1_dataframe$renyi.long1.Grouping, levels = c('Control farm','Bare fence',"One year old","Two year old","Three year old","Old"))

pl_renyi_across <- ggplot(data=renyi.long1_dataframe, aes(x=renyi.long1.Scales, y=renyi.long1.Diversity, ymax=renyi.long1.UPR, ymin=renyi.long1.LWR)) + 
    scale_x_discrete() +
    scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
    geom_line(data=renyi.long1_dataframe, 
              aes(x=renyi.long1.Obs, colour=Boundary_descriptor), 
              linewidth=2) +
    geom_point(data=subset(renyi.long1_dataframe, renyi.long1.labelit==TRUE), 
        aes(colour=Boundary_descriptor, shape=Boundary_descriptor), 
        size=5) +
    geom_ribbon(data=renyi.long1_dataframe, 
                aes(x=renyi.long1.Obs, colour=Boundary_descriptor, fill=after_scale(alpha(colour, 0.1))), 
                show.legend=FALSE) + 
     scale_colour_manual(values=col_bound)+
    theme_bw() +
    #scale_colour_npg() +
    labs(x=expression(alpha), y = "Diversity", colour = "Boundary type", shape = "Boundary type")
pl_renyi_across

  
#2022/23
meta_data_bior_season<-meta_data_bior %>% filter(Season=='2022/23')
meta_data_bior_season %>% select(Boundary_descriptor) %>% group_by(Boundary_descriptor) %>% summarise(count = n())
meta_data_bior_season$Boundary_descriptorf<-factor(meta_data_bior_season$Boundary_descriptor, levels = c('Control farm','Bare fence',"Two year old","Old"))
counts_wide_filt <- counts_wide[rownames(counts_wide) %in% meta_data_bior_season$unique_sample, ]
counts_wide_filt <- counts_wide_filt[,colSums(counts_wide_filt) > 0]
Renyi.1<-renyicomp(counts_wide_filt,y=meta_data_bior_season,factor='Boundary_descriptorf', 
  scales=c(0, 0.25, 0.5, 1, 2, 4, 8, Inf), permutations=1000, plotit=F)
renyi.long1 <- renyicomp.long(Renyi.1, label.freq=1)
renyi.long1_dataframe<-data.frame(renyi.long1$labelit,renyi.long1$Scales,renyi.long1$Obs,renyi.long1$Diversity,renyi.long1$UPR,renyi.long1$LWR,renyi.long1$Grouping)
renyi.long1_dataframe$Boundary_descriptor<-factor(renyi.long1_dataframe$renyi.long1.Grouping, levels = c('Control farm','Bare fence',"Two year old","Old"))
col_bound_cut<-c('chocolate','#FDE725FF',"#21908CFF","#440154FF")


pl_renyi_across <- ggplot(data=renyi.long1_dataframe, aes(x=renyi.long1.Scales, y=renyi.long1.Diversity, ymax=renyi.long1.UPR, ymin=renyi.long1.LWR)) + 
    scale_x_discrete() +
    scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
    geom_line(data=renyi.long1_dataframe, 
              aes(x=renyi.long1.Obs, colour=Boundary_descriptor), 
              linewidth=2) +
    geom_point(data=subset(renyi.long1_dataframe, renyi.long1.labelit==TRUE), 
        aes(colour=Boundary_descriptor, shape=Boundary_descriptor), 
        size=5) +
    geom_ribbon(data=renyi.long1_dataframe, 
                aes(x=renyi.long1.Obs, colour=Boundary_descriptor, fill=after_scale(alpha(colour, 0.1))), 
                show.legend=FALSE) + 
     scale_colour_manual(values=col_bound_cut)+
    theme_bw() +
    #scale_colour_npg() +
    labs(x=expression(alpha), y = "Diversity", colour = "Boundary type", shape = "Boundary type")
pl_renyi_across


#2021/22
meta_data_bior_season<-meta_data_bior %>% filter(Season=='2021/22')
meta_data_bior_season %>% select(Boundary_descriptor) %>% group_by(Boundary_descriptor) %>% summarise(count = n())
meta_data_bior_season$Boundary_descriptorf<-factor(meta_data_bior_season$Boundary_descriptor, levels = c('Control farm','Bare fence',"One year old","Old"))
counts_wide_filt <- counts_wide[rownames(counts_wide) %in% meta_data_bior_season$unique_sample, ]
counts_wide_filt <- counts_wide_filt[,colSums(counts_wide_filt) > 0]
Renyi.1<-renyicomp(counts_wide_filt,y=meta_data_bior_season,factor='Boundary_descriptorf', 
  scales=c(0, 0.25, 0.5, 1, 2, 4, 8, Inf), permutations=1000, plotit=F)
renyi.long1 <- renyicomp.long(Renyi.1, label.freq=1)
renyi.long1_dataframe<-data.frame(renyi.long1$labelit,renyi.long1$Scales,renyi.long1$Obs,renyi.long1$Diversity,renyi.long1$UPR,renyi.long1$LWR,renyi.long1$Grouping)
renyi.long1_dataframe$Boundary_descriptor<-factor(renyi.long1_dataframe$renyi.long1.Grouping, levels = c('Control farm','Bare fence',"One year old","Old"))
col_bound_cut<-c('chocolate','#FDE725FF',"#5DC863FF","#440154FF")

pl_renyi_across <- ggplot(data=renyi.long1_dataframe, aes(x=renyi.long1.Scales, y=renyi.long1.Diversity, ymax=renyi.long1.UPR, ymin=renyi.long1.LWR)) + 
    scale_x_discrete() +
    scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
    geom_line(data=renyi.long1_dataframe, 
              aes(x=renyi.long1.Obs, colour=Boundary_descriptor), 
              linewidth=2) +
    geom_point(data=subset(renyi.long1_dataframe, renyi.long1.labelit==TRUE), 
        aes(colour=Boundary_descriptor, shape=Boundary_descriptor), 
        size=5) +
    geom_ribbon(data=renyi.long1_dataframe, 
                aes(x=renyi.long1.Obs, colour=Boundary_descriptor, fill=after_scale(alpha(colour, 0.1))), 
                show.legend=FALSE) + 
     scale_colour_manual(values=col_bound_cut)+
    theme_bw() +
    #scale_colour_npg() +
    labs(x=expression(alpha), y = "Diversity", colour = "Boundary type", shape = "Boundary type")
pl_renyi_across

#hmmm pattern holds across seasons, I thought maybe that it would be one season throwing everyone out

#just having a look at the accum results:
Accum.1 <- accumcomp(counts_wide,y=meta_data_bior,factor='Boundary_descriptorf', 
    method='exact', conditioned=FALSE, plotit=FALSE)
accum.long1 <- accumcomp.long(Accum.1, ci=NA, label.freq=5)
accum.long1_dataframe<-data.frame(accum.long1$Grouping,accum.long1$Obs,accum.long1$Sites,accum.long1$Richness,accum.long1$LWR,accum.long1$UPR,accum.long1$labelit)
accum.long1_dataframe$Boundary_descriptor<-factor(accum.long1_dataframe$accum.long1.Grouping, levels = c('Control farm','Bare fence',"One year old","Two year old","Three year old","Old"))


plotgg1 <- ggplot(data=accum.long1_dataframe, aes(x = accum.long1.Sites, y = accum.long1.Richness, ymax = accum.long1.UPR, ymin = accum.long1.LWR)) + 
    scale_x_continuous(expand=c(0, 1), sec.axis = dup_axis(labels=NULL, name=NULL)) +
    scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
    geom_line(aes(colour=Boundary_descriptor), linewidth=2) +
    geom_point(data=subset(accum.long1_dataframe, accum.long1.labelit==TRUE), 
               aes(colour=Boundary_descriptor, shape=Boundary_descriptor), size=5) +
    geom_ribbon(aes(colour=Boundary_descriptor), alpha=0.2, show.legend=FALSE) + 
     scale_colour_manual(values=col_bound)+
    theme_bw() +
    labs(x = "Sites", y = "Species", colour = "Boundary type", shape = "Boundary type")
plotgg1
#

```


## Beta Diversity
Beta diversity analysis
Adonis2 works with nesting via a strata factor that blocks permutations to within groups however if you add both Date and Farm it doesnt have enough space to shuffle things around within the groups.

This analysis is important but Im finding it the most difficult as it makes a large difference how I structure the data (the other ones it doesnt matter what I do the answer is the same). In the other diversity tests, it really doesnt make much difference to the overall result but here it does. 


```{r beta_alpha, echo=T}
#beta diversity
#cant run an offset in adonis so flower count going in as a variable
perm<-adonis2(counts_wide ~ Boundary_descriptor + `Flower count`+Season, data = meta_data)
perm

#we need to restrict permutations around nesting:
####################################
#so if site and date were irreverent I could just do
#strata for season
perm<-adonis2(counts_wide ~ Boundary_descriptor + `Flower count`, data = meta_data, strata = meta_data$Season)
perm
pairwise_adonis<-pairwise.adonis2(counts_wide ~ Boundary_descriptor + `Flower count`,data=meta_data_bior,strata='Season')
pairwise_adonis

#but you know not in my ideal world and I cant work out the best way to get both date and farm in. 
#adonis can only take one strata value so either combine into one
#season farm and month split
meta_data$site_day_seas <- interaction(meta_data$Farm_name, meta_data$month_split,meta_data$Season, drop = TRUE)
perm<-adonis2(counts_wide ~ Boundary_descriptor +`Flower count`, data = meta_data, strata = meta_data$site_day_seas)
perm
#so we can say there is a significant impact of boundary 

#but the pairwise adhoc afterwards doesnt work as their is no movement for strata between groups
#so dropping back to season and month
meta_data_bior$month_season <- interaction(meta_data_bior$Collection_month, meta_data_bior$Season, drop = TRUE)
pairwise_adonis<-NULL
pairwise_adonis<-pairwise.adonis2(counts_wide ~ Boundary_descriptor +`Flower count`,data=meta_data_bior, strata='month_season')
pairwise_adonis

#if I wanted to include farm Im not sure that I can with season, there is too many that are singles
#I've tried every permutation and there just isnt enough freedom in the strata system to get those in
#season has by far the biggest impact so season/month is the best we can do maybe?

```

## NMDS
NMDS plot

```{r nmds_alpha, echo=T}
#NMDS is struggling to find a repeated solution. I think this is due to the rare species
#https://stats.stackexchange.com/questions/649643/best-solution-was-not-repeated-no-reliable-result-for-metamds

pk_NMDS <- metaMDS(counts_wide,trymax=200,distance='bray')

pk_NMDS
#stress is a little high (ideally <0.2)

stressplot(pk_NMDS)

plot(pk_NMDS)

plot_df <- scores(pk_NMDS, display = "sites") %>% 
  as.data.frame() %>% 
  rownames_to_column("site") %>% 
  full_join(meta_data, by = c("site" = "unique_sample"))

plot_df$Boundary_descriptor <- factor(plot_df$Boundary_descriptor,levels=c('Control farm','Bare fence','One year old','Two year old','Three year old','Old'),ordered = T)


plot_nmds <- ggplot(plot_df, aes(x = NMDS1, y = NMDS2, color = Boundary_descriptor)) +
  geom_point(size = 2, alpha = 0.8) +
  stat_ellipse(linetype = 2, linewidth = 0.5) +
  theme_bw()+
    scale_colour_manual(values=col_bound)+
  labs(title = "NMDS",
       colour = "Boundary type")
plot_nmds
```


## RDA
An alternative to the MDS that would align well with the adonis2 beta diversity analysis would be a rda (constrained PCA - PCA on predicted variables) and using the model. 
For the species lines note that I double the length here for plotting


```{r rda_mds, echo=T,warning = FALSE}
counts_wide.Hellinger <- disttransform(counts_wide, method='hellinger')
#add conditions here for random effects:
Ordination.model2 <- rda(counts_wide.Hellinger ~ Boundary_descriptorf +  offset(log(`Flower count`)) + Condition(Season)+Condition(Farm_name)+Condition(month_split), data=meta_data_bior, scaling="species")

plot2 <- ordiplot(Ordination.model2, choices=c(1,2))
sites.long2 <- sites.long(plot2, env.data=meta_data_bior)
species.long2 <- species.long(plot2)
axis.long2 <- axis.long(Ordination.model2, choices=c(1, 2))
sites.long2$Boundary_descriptor <- factor(sites.long2$Boundary_descriptor,levels=c('Control farm','Bare fence','One year old','Two year old','Three year old','Old'),ordered = T)

#just dots
plotgg2 <- ggplot() + 
    geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
    geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
    xlab(axis.long2[1, "label"]) +
    ylab(axis.long2[2, "label"]) +  
    scale_x_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
    scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +    
    geom_point(data=sites.long2, 
               aes(x=axis1, y=axis2, colour=Boundary_descriptor), 
               size=3) +
   scale_colour_manual(values=col_bound)+
    theme_bw()

plotgg2

#95% circles
plotgg2 <- ggplot() + 
    geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
    geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
    xlab(axis.long2[1, "label"]) +
    ylab(axis.long2[2, "label"]) +  
    scale_x_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
    scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +    
    geom_point(data=sites.long2, 
               aes(x=axis1, y=axis2, colour=Boundary_descriptor), 
               size=3) +
  stat_ellipse(data=sites.long2, 
                      aes(x=axis1, y=axis2, colour=Boundary_descriptor, fill=Boundary_descriptor),
type='norm',level=0.95,geom='polygon',alpha=0.1,show.legend=F) +
    scale_colour_manual(values=col_bound)+
     scale_fill_manual(values=col_bound)+
   theme_bw()

plotgg2

#add in species arrows
spec.envfit <- envfit(plot2, env=counts_wide.Hellinger)
spec.data.envfit <- data.frame(r=spec.envfit$vectors$r, p=spec.envfit$vectors$pvals)
species.long2 <- species.long(plot2, spec.data=spec.data.envfit)
#species that explain 10% of variation
species.long3 <- species.long2[species.long2$r >= 0.1, ]
#species.long3
ordiArrowMul(spec.envfit) #close to 2

plotgg2 <- ggplot() + 
    geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
    geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
    xlab(axis.long2[1, "label"]) +
    ylab(axis.long2[2, "label"]) +  
    scale_x_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
    scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +    
    geom_point(data=sites.long2, 
               aes(x=axis1, y=axis2, colour=Boundary_descriptor), 
               size=3) +
   scale_colour_manual(values=col_bound)+
    geom_segment(data=species.long3, 
                 aes(x=0, y=0, xend=axis1*2, yend=axis2*4), 
                 colour="black", size=0.7, arrow=arrow()) +
    geom_text_repel(data=species.long3, 
                    aes(x=axis1*2, y=axis2*2, label=labels),
                    colour="black") +
    theme_bw()

plotgg2

#species and circles
plotgg2 <- ggplot() + 
    geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
    geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
    xlab(axis.long2[1, "label"]) +
    ylab(axis.long2[2, "label"]) +  
    scale_x_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
    scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +    
    geom_point(data=sites.long2, 
               aes(x=axis1, y=axis2, colour=Boundary_descriptor), 
               size=3) +
  stat_ellipse(data=sites.long2, 
                      aes(x=axis1, y=axis2, colour=Boundary_descriptor, fill=Boundary_descriptor),
type='norm',level=0.95,geom='polygon',alpha=0.1,show.legend=F) +
    scale_colour_manual(values=col_bound)+
     scale_fill_manual(values=col_bound)+
    geom_segment(data=species.long3, 
                 aes(x=0, y=0, xend=axis1*2, yend=axis2*2), 
                 colour="black", size=0.7, arrow=arrow()) +
    geom_text_repel(data=species.long3, 
                    aes(x=axis1*2, y=axis2*2, label=labels),
                    colour="black") +
    theme_bw()

plotgg2

#alternative would be ddRDA or capscale on bray distance (PCoA) which might be better for us for differences between communities
cap <- capscale(counts_wide ~ Boundary_descriptorf + offset(log(`Flower count`)) + Condition(Season)+Condition(Farm_name)+Condition(month_split), data = meta_data_bior, distance = "bray")
#cap <- capscale(counts_wide.Hellinger ~ Boundary_descriptorf + offset(log(`Flower count`)) + Condition(Season)+Condition(Farm_name)+Condition(month_split), data = meta_data_bior, distance = "bray")

anova(cap, by = "terms")
plot(cap)
site_scores <- scores(cap, display = "sites")
species_scores <- scores(cap, display = "species")
centroids <- scores(cap, display = "cn")


df_sites <- as.data.frame(site_scores)
df_sites$unique_sample <- rownames(df_sites)
capscale_df <- left_join(df_sites,meta_data)

capscale_df$Boundary_descriptor <- factor(capscale_df$Boundary_descriptor,levels=c('Control farm','Bare fence','One year old','Two year old','Three year old','Old'),ordered = T)

eig_vals <- cap$CCA$eig
prop_var <- eig_vals / sum(eig_vals)

xlab <- paste0("CAP1 (", round(prop_var[1] * 100, 1), "%)")
ylab <- paste0("CAP2 (", round(prop_var[2] * 100, 1), "%)")

#base plot
ggplot(capscale_df, aes(x = CAP1, y = CAP2, color = Boundary_descriptor)) +
  geom_point(size = 3) +
 scale_colour_manual(values=col_bound)+
  labs(x = xlab, y = ylab, color='Boundary type') +
  theme_bw()

#95% circles  
ggplot() +
  geom_point(data=capscale_df, aes(x = CAP1, y = CAP2, color = Boundary_descriptor),size = 3) +
  scale_colour_manual(values=col_bound)+
  scale_fill_manual(values=col_bound)+
  labs(x = xlab, y = ylab, color='Boundary type') +
  geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
  geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
  stat_ellipse(data=capscale_df, 
  aes(x = CAP1, y = CAP2, color = Boundary_descriptor,fill = Boundary_descriptor), 
  type='norm',level=0.95,geom='polygon',alpha=0.1,show.legend=F)+
  theme_bw()

#note species scores are calculated using euclidean distance in capscale which we assume correlates to brays
#so probably dump the species on in supplementary and say combined with standard RDA shows driven by lassio
# Calculate arrow length (distance from origin) to equate to species explaining most variation
#top species
sp_len <- sqrt(rowSums(species_scores^2))

# Rank species by length
head(sort(sp_len, decreasing = TRUE), 10)   # top 10 species for plot
keep <- names(sp_len[sp_len > quantile(sp_len, 0.9)])  # top 10% longest arrows
sp_filt <- species_scores[keep, ]

#species in plot      
ggplot() +
  geom_point(data=capscale_df, aes(x = CAP1, y = CAP2, color = Boundary_descriptor),size = 3) +
  scale_colour_manual(values=col_bound)+
  labs(x = xlab, y = ylab, color='Boundary type') +
  geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
  geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
  geom_segment(data=sp_filt, 
     aes(x=0, y=0, xend=CAP1*2, yend=CAP2*2), 
     colour="black", size=0.7, arrow=arrow()) +
  geom_text_repel(data=sp_filt, 
    aes(x=CAP1*2, y=CAP2*2, label=rownames(sp_filt)),
    colour="black") +
  theme_bw()
      
#circles      
ggplot() +
  geom_point(data=capscale_df, aes(x = CAP1, y = CAP2, color = Boundary_descriptor),size = 3) +
  scale_colour_manual(values=col_bound)+
  scale_fill_manual(values=col_bound)+
  labs(x = xlab, y = ylab, color='Boundary type') +
  geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
  geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
  geom_segment(data=sp_filt, 
     aes(x=0, y=0, xend=CAP1*2, yend=CAP2*2), 
     colour="black", size=0.7, arrow=arrow()) +
  geom_text_repel(data=sp_filt, 
    aes(x=CAP1*2, y=CAP2*2, label=rownames(sp_filt)),
    colour="black") +
  stat_ellipse(data=capscale_df, 
  aes(x = CAP1, y = CAP2, color = Boundary_descriptor,fill = Boundary_descriptor), 
  type='norm',level=0.95,geom='polygon',alpha=0.1,show.legend=F)+
  theme_bw()

#I think ddRDA is probably slightly better than capscale from what I've read but they do a similar thing 

bound_dbrda <- dbrda(counts_wide ~ Boundary_descriptorf + offset(log(`Flower count`)) + Condition(Season)+Condition(Farm_name)+Condition(month_split), data = meta_data_bior, distance = "bray")
#bound_dbrda <- dbrda(counts_wide.Hellinger ~ Boundary_descriptorf + offset(log(`Flower count`)) + Condition(Season)+Condition(Farm_name)+Condition(month_split), data = meta_data_bior, distance = "bray")

bound_dbrda_summary<-summary(bound_dbrda)
plot(bound_dbrda)
anova(bound_dbrda, by = "terms")

site_scores <- scores(bound_dbrda, display = "sites")
centroids <- scores(bound_dbrda, display = "cn")


df_sites <- as.data.frame(site_scores)
df_sites$unique_sample <- rownames(df_sites)
bound_dbrda_df <- left_join(df_sites,meta_data)

bound_dbrda_df$Boundary_descriptor <- factor(capscale_df$Boundary_descriptor,levels=c('Control farm','Bare fence','One year old','Two year old','Three year old','Old'),ordered = T)

eig_vals <- bound_dbrda$CCA$eig
prop_var <- eig_vals / sum(eig_vals)

xlab <- paste0("dbRDA1 (", round(prop_var[1] * 100, 1), "%)")
ylab <- paste0("dbRDA2 (", round(prop_var[2] * 100, 1), "%)")

ggplot(bound_dbrda_df, aes(x = dbRDA1, y = dbRDA2, color = Boundary_descriptor)) +
  geom_point(size = 3) +
 scale_colour_manual(values=col_bound)+
  labs(x = xlab, y = ylab, color='Boundary type') +
  theme_bw()

#with circles
#95% circles  
ggplot() +
  geom_point(data=bound_dbrda_df, aes(x = dbRDA1, y = dbRDA2, color = Boundary_descriptor),size = 3) +
  scale_colour_manual(values=col_bound)+
  scale_fill_manual(values=col_bound)+
  labs(x = xlab, y = ylab, color='Boundary type') +
  geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
  geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
  stat_ellipse(data=bound_dbrda_df, aes(x = dbRDA1, y = dbRDA2, color = Boundary_descriptor,fill=Boundary_descriptor), 
  type='norm',level=0.95,geom='polygon',alpha=0.1,show.legend=F)+
  theme_bw()

#add in species arrows
#rdRDA doesnt add species scores in automatically (due to them being calculated from the eucludian distance)
#using hellinger transformed here
sppscores(bound_dbrda)<-counts_wide.Hellinger

# Calculate arrow length (distance from origin) to equate to species explaining most variation
#top species
species_scores <- scores(bound_dbrda, display = "species")

sp_len <- sqrt(rowSums(species_scores^2))

# Rank species by length
head(sort(sp_len, decreasing = TRUE), 10)   # top 10 species for plot
keep <- names(sp_len[sp_len > quantile(sp_len, 0.9)])  # top 10% longest arrows
sp_filt <- species_scores[keep, ]

#species in plot      
ggplot() +
  geom_point(data=bound_dbrda_df, aes(x = dbRDA1, y = dbRDA2, color = Boundary_descriptor),size = 3) +
  scale_colour_manual(values=col_bound)+
  labs(x = xlab, y = ylab, color='Boundary type') +
  geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
  geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
  geom_segment(data=sp_filt, 
     aes(x=0, y=0, xend=dbRDA1*2, yend=dbRDA2*2), 
     colour="black", size=0.7, arrow=arrow()) +
  geom_text_repel(data=sp_filt, 
    aes(x=dbRDA1*2, y=dbRDA2*2, label=rownames(sp_filt)),
    colour="black") +
  theme_bw()

#add in species arrows & eclipses
ggplot() +
  geom_point(data=bound_dbrda_df, aes(x = dbRDA1, y = dbRDA2, color = Boundary_descriptor),size = 3) +
  scale_colour_manual(values=col_bound)+
  scale_fill_manual(values=col_bound)+
  labs(x = xlab, y = ylab, color='Boundary type') +
  geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
  geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
  geom_segment(data=sp_filt, 
     aes(x=0, y=0, xend=dbRDA1*2, yend=dbRDA2*2), 
     colour="black", size=0.7, arrow=arrow()) +
  stat_ellipse(data=bound_dbrda_df, aes(x = dbRDA1, y = dbRDA2, color = Boundary_descriptor,fill=Boundary_descriptor), 
  type='norm',level=0.95,geom='polygon',alpha=0.1,show.legend=F)+
  geom_text_repel(data=sp_filt, 
    aes(x=dbRDA1*2, y=dbRDA2*2, label=rownames(sp_filt)),
    colour="black") +
  theme_bw()

#given the issues with pairwise adonis tests we could test the mean separation on the ordination axes with emmeans (maybe?)
# Extract site scores from dbRDA
site_scores <- scores(bound_dbrda, display = "sites")
site_scores_df <- data.frame(site_scores, Boundary_descriptor = meta_data$Boundary_descriptor)

# Fit linear models per axis
lm1 <- lm(dbRDA1 ~ Boundary_descriptor, data = site_scores_df)
lm2 <- lm(dbRDA2 ~ Boundary_descriptor, data =site_scores_df)

# Pairwise comparisons
library(emmeans)
emmeans(lm1, pairwise ~ Boundary_descriptor, adjust = "tukey")
emmeans(lm2, pairwise ~ Boundary_descriptor, adjust = "tukey")
#

```

## iNEXT

iNEXT (iNterpolation and EXTrapolation) is an R package, available in CRAN and Github, for rarefaction and extrapolation of species diversity (Hill numbers). Its a useful package to use if you want to make any predictions about sampling size or diversity. 

### Hill numbers

Hill numbers are a set of diversity indices used in ecology and biodiversity assessments to quantify the richness and evenness of species in a given community. These indices were developed by Michael Hill and are widely used to summarize various aspects of biodiversity. Hill numbers are particularly useful because they provide a way to focus on different facets of diversity, depending on the specific questions or goals of a study. For a thorough introduction to Hill numbers check out: Alberdi & Gilbert


There are several Hill numbers, but they all share a common formula that relies on an exponent, denoted as "q." The choice of the exponent determines which aspect of diversity is emphasized:

- Richness (q = 0): Hill number with q = 0 is also known as the "species richness" or "taxonomic distinctness" index. It simply counts the number of unique species in a community without considering their abundances. This index is valuable when you want to emphasize the presence or absence of species but not their relative abundance.

- Exponential Shannon entropy (q = 1): This index corresponds to the Shannon-Wiener diversity index (commonly referred to as Shannon entropy) when q is set to 1. It considers both the richness and the evenness of species abundances. It is often used when you want to give equal importance to species richness and evenness.

- Inverse Simpson index (q = 2): The Inverse Simpson index corresponds to the Simpson diversity index when q is set to 2. This index emphasizes the dominance of the most abundant species in a community. It is often used when you want to give more weight to the presence of dominant species.

- Higher Hill numbers (q > 2): These Hill numbers give increasing weight to the most abundant species, placing more emphasis on the dominant species in the community. The larger the value of q, the more sensitive the index becomes to the presence of dominant species.

Hill numbers offer a flexible framework for assessing biodiversity because they allow you to choose the appropriate value of q based on the specific ecological question or management objective you are addressing. For example:

- If you want to measure the overall diversity of a community while considering both species richness and evenness, you might use q = 1.

- If you are concerned about the impact of a few highly dominant species in an ecosystem, you might use q = 2 or a higher value to focus on those dominant species.

#iNEXT package
iNEXT is a neat package that can calculate Hill numbers, enabling us to investigate alpha diversity accumulation curves (similar to species accumulation curves). These graphs can help us determine if sufficient sampling was conducted at each site, based on the plateauing of the curves, similar to the rarefaction curves. But iNEXT.3D goes beyond rarefraction and extrapolates alpha diversity numbers beyond the number of samples collected (default is double sample number). Its written by Anne Chao and you can find her papers here: https://sites.google.com/view/chao-lab-website/home

### Running iNEXT

```{r inext, echo=T,warning = FALSE}
#innext uses a weird list of lists, this is a annoying format to get into
#im using raw incident data as the input (thus it is made binary at one point in the code)

#generate a list of the unique sample locations
categories <- unique(data$Boundary_descriptor)
categories

#create empty list
split_list <- list()
#fill empty list with a presence absence OTU from each location

data<-data %>% mutate(unique_ob=paste(Boundary_descriptor,Farm_name,Season)) 

for (category in categories) {
  subset_data <- data %>% filter(Boundary_descriptor == category)
  test<-subset_data %>% ungroup() %>% dplyr::select(unique_ob,`New Inseck key name`,`Sum of Insects/100 flowers`)
  test   <- test %>%  spread(key = unique_ob, value = `Sum of Insects/100 flowers`)
  test<-test %>% remove_rownames %>% column_to_rownames(var="New Inseck key name")
  test<-test %>% mutate(across(everything(), ~ ifelse(. > 0, 1, 0)))
  split_list[[category]]<-as(test,'matrix')
} 

matrix_list <- list(data = list())
for (category in categories) {
  otu_table <- as(split_list[[category]], "matrix")
  matrix_list[["data"]][[category]] <- otu_table
}

#here Im setting the endpoint as NULL which is double the input sample size (this is probably most robust)
out.raw <- iNEXT3D(data = matrix_list$data, diversity = 'TD', q = c(0, 1, 2), datatype = 'incidence_raw', nboot = 50,endpoint=NULL)

#there is various ways to look at these plots:
#within a sample
ggiNEXT3D(out.raw, type = 1, facet.var = 'Assemblage') + facet_wrap(~Assemblage, nrow = 3)
#across samples
ggiNEXT3D(out.raw, type = 1, facet.var = "Order.q")
#sample completeness
ggiNEXT3D(out.raw, type = 2, facet.var = "Order.q", color.var = "Assemblage")

#using innext diversity estimates in a model


```

# Alternative plots for iNEXT

Here i have set the endpoint as 50 - I doubt this is robust for the 1,2,3 year plots but makes the plots even. 'q1' in particular looks squiffy at 50.

```{r inext2, echo=T,warning = FALSE}
out.raw <- iNEXT3D(data = matrix_list$data, diversity = 'TD', q = c(0, 1, 2), datatype = 'incidence_raw', nboot = 50,endpoint=50)

#there is various ways to look at these plots:
#within a sample
ggiNEXT3D(out.raw, type = 1, facet.var = 'Assemblage') + facet_wrap(~Assemblage, nrow = 3)
#across samples
ggiNEXT3D(out.raw, type = 1, facet.var = "Order.q")
#sample completeness
ggiNEXT3D(out.raw, type = 2, facet.var = "Order.q", color.var = "Assemblage")

```



